CrypTool Script \\ Mathematics and Cryptography

### 2002-01-18 script-gb/script-gb.tex
no message

\documentclass[a4paper,11pt]{article}
% Da man keine Unterst"utzung der deutschen Sprache in diesem Fall braucht!!
% \usepackage[german]{babel}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{makeidx}
\usepackage{graphicx}
\makeindex
\hypersetup{colorlinks=true}

@@

\begin{document}

@@

You will receive information about the principles of symmetrical and
asymmetrical encryption. A large section of the script is dedicated to the
fascinating topic of prime numbers. With the use of numerous examples, the elementary number
theory and modular arithmetic have been introduced and have been applyed in an exemplary 
manner in the RSA procedure.  After this, one receives insight into the mathematical ideas
behind modern cryptography.

@@

The {\em authors} Bernhard Esslinger, Bartol Filipovic, Henrik Koy and Roger Oyono
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Gie{\ss}en und Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

@@

\subsubsection{Asymmetrc encryption}

@@

Although there are only $92$ natural chemical elements, the number of prime
numbers is unlimited. Even the Greek, \index{Euclid} Euclid\footnote{
Euclid, a Greek mathematician of 4th and 3rd century before Christ. He worked at the
egyptian academy of Alexandria and wrote ''The Elements'', the most well known systematical textbook
of the Greek mathematics.} knew this in the
third century before Christ.
\begin{theorem}[Euclid]\footnote{
The common usage of the term does not denote Euclid as the inventor of the theorem rather;
the true inventor is merely not as prominent. he theorem has already been distinguished
and proven in Euclid’s Elements (Book IX, s. 20). The phraseology is remarkable due to 
the fact that the word infinite is not used. The text reads as following
$$
O\acute{\iota}~\pi\varrho\tilde{\omega}\tau o \iota~\grave{\alpha}\varrho\iota\vartheta\mu o\grave{\iota}~
\pi\lambda\varepsilon\acute{\iota}o \upsilon\varsigma~\varepsilon\grave{\iota}\sigma\grave{\iota}~
\pi\alpha\nu\tau\grave{o}\varsigma~\tau o \tilde{\upsilon}~
\pi\varrho o \tau\varepsilon\vartheta\acute{\varepsilon}\nu\tau o \varsigma~
\pi\lambda\acute{\eta}\vartheta\ o \upsilon\varsigma~
\pi\varrho\acute{\omega}\tau\omega\nu~
\grave{\alpha}\varrho\iota\vartheta\mu\tilde{\omega}\nu,
$$
the English translation of which is: the prime numbers are more than any previously existing amount
of prime numbers.}
The sequence of prime numbers does not discontinue.
Therefore, the quantity of prime numbers is infinite.
\end{theorem}
His proof that there is an infinite number of
primes is still considered to be a brilliant mathematical consideration and
conclusion today (proof by contradiction). He assumed that there is only a
finite number of primes and therefore a largest prime number. Based on this
assumption, he drew logical conclusions until he obtained an obvious
contradiction. This meant that something must be wrong. As there were no
mistakes in the chain of conclusions, it could only be the assumption that was
wrong. Therefore, there must be an infinite number of primes!

@@

Here we make mention of yet another fact which is initially somewhat surprising. 
Namely, in the prime numbers sequence $p_1, p_2, \cdots,$ gaps between prime numbers can have
an individually determined length $n$. It is undeniable that under the $n$
succession of natural numbers
$$(n+1)!+2,\cdots, (n+1)!+(n+1),
$$
none of them is a prime number since in order, the numbers $2,3,\cdots,(n+1)$  
are comprised respectively as real divisors. 
Nevertheless $n!$ means the product of the first $n$ natural numbers therefore 
$ n!= n*(n-1)*\cdots *2*1.$

@@

The 39th Mersenne prime, called M-39, $$2^{13,466,917}-1,$$ was found at December 6, 2001 – more exactly,
the verification of this number, found at November 14, 2001 by the Canadian student Michael Cameron,
was successfully finished. This number has about 4 million decimal digits (exactly 4,053,946 digits).
Trying only to print this number $(924947738006701322247758 \cdots 1130073855470256259071)$ in the Financial Times
would need around 200 pages.

@@

  \item {\em Fermat numbers}\footnote{
The Fermat prime numbers play a role in circel division. As proven by Gauss, a regular $p$-edge
can only be constructed with the use of a compass and a ruler, when $p$ is a Fermat prime number.}
\index{Numbers!Fermat numbers}\index{Fermat!Fermat
number}  $F(n) = 2^{2^n} + 1$:\\
As mentioned above, Fermat wrote to Mersenne regarding this assumption.
Surprisingly he would have been able obtain a positive result using the negative
prime number test for $n=5$ based on his small theorem.
$$
\begin{array}{lll}
F(0) = 2^{2^0} + 1  = 2^1 + 1 & = 3 &   \mapsto {\rm ~prime}  \\
F(1) = 2^{2^1} + 1  = 2^2 + 1 & = 5 &   \mapsto {\rm ~prime}  \\
F(2) = 2^{2^2} + 1  = 2^4 + 1 & = 17 &  \mapsto {\rm ~prime}  \\
F(3) = 2^{2^3} + 1  = 2^8 + 1 & = 257 & \mapsto {\rm ~prime}  \\
F(4) = 2^{2^4} + 1  = 2^{16} + 1 &  = 65,537 &  \mapsto {\rm ~prime}  \\
F(5) = 2^{2^5} + 1  = 2^{32} + 1 &  = 4 ,294, 967, 297 = 641 \cdot
6,700,417 &  \mapsto {\rm ~NOT~prime} !
\end{array} 
$$
  \item Carmichael numbers: see above.
  \item Pseudoprime numbers: see above.
  \item Strong pseudoprime numbers: see above.
  \item Idea based on \hyperlink{euclid}{Euclid's proof} (infinite many prime
numbers) $p_1 \cdot p_2 \cdots p_n +1$:

@@

$e_9, \cdots, e_{17}$ are also composite, which means that this formula is not
particularly useful. Comment: However, what is special about all these numbers
is that any pair of them does not have a common factor other than $1$. They are
therefore \index{Prime numbers!relative} {\em relatively prime }.
  \item $f(n) = n^2 + n + 41$:\\
  This sequence starts off very {\em promisingly},   but is far from being a
proof.

@@

This is the \index{Prime number theorem} \textbf{prime number theorem}. It was
put forward by Legendre (1752-1833) and Gau\ss\ (1777-1855) but not proved until
over 100 years later.

@@

\begin{itemize}
  \item The German mathematician \index{Goldbach Christian} Christian Goldbach
(1690-1764) conjectured:
    Every even natural number greater than 2 can be represented as the sum of
two prime numbers.
    Computers have verified\footnote{
It is generally accepted today, that the Goldbach Conjecture is true, i.e. valid for all prime numbers.
In 1999, mathematician J\"org Richtstein from the computer sciences institute at the University
of Giessen, studied even numbers up to 400 billion and found no contradictory example (see
\href{http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}{\tt http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}).\index{Richstein 1999}
Nevertheless, this does not provide us with general proof.

The fact that despite all efforts, Goldbachs' conjecture has to date not been proven leads
one to believe that: since the pioneer work of the Austrian mathematician Kurt Gödel is well-known,
not every true mathematical theorem is provable (see
\href{http://www.mathematik.ch/mathematiker/goedel.html}{http://www.mathematik.ch/mathematiker/goedel.html}).
Perhaps Goldbachs' Conjecture was correct, despite that however,
proof will never be found. Conversely, that will presumably also remain unproven.}
the Goldbach conjecture for all even numbers up to
$4*10^{14}$ but no general proof has yet been found\footnote{
The English publisher {\em Faber} and the American publisher {\em Bloomsbury} issued in 2000
the 1992 published book ''Uncle Petros and Goldbach's Conjecture'' by Apostolos Doxiadis.
It's the story of an old maths professor who fails to prove a more than 250 old puzzle. To
stimulate the sell rates the English and American publisher compete a prize of 1 million USD, 
is someone proves the conjecture -- to be published till 2004 by an well-known mathematical journal.

Susprisingly only British and American citizens are allowed to participate.

The theorem closest to Goldbach's conjecture was proven 1966 by Chen Jing-Run in a hard to understand fashion:
Each even integer bigger than $2$ is the sum of one prime and of the product of two primes. E.g.: $20=5+3*5.$

Most of the research about the Goldbach conjecture is collected in the book: ''Goldbach Conjecture'', ed. Wang Yuan, 1984, World scientific
Series in Pure Maths, Vol. 4.
}.
  \item The German \index{Riemann Bernhard} mathematician Bernhard Riemann
(1826-1866) put forward a formula for the distribution of primes that would
further improve the estimate. However, this has neither been proved nor
contradicted (?) so far.
\end{itemize}

@@

Twin primes are prime numbers whose difference is 2. Examples include 5 and 7 or
101 and 103. Triplet primes, however, only occur once: 3, 5, 7.  For all other
sets of three consecutive uneven numbers, one of them is always divisible by 3
and thus not a prime.
\begin{itemize}
  \item The number of twin primes is an open question: infinite or limited
number?
  The largest twin primes known today are $1,693,965 \cdot 2^{66,443} \pm 1.$
  \item Does a formula exist for calculating the number of twin primes per
interval?
  \item The above proof of the function  $f: N \rightarrow Z$ with $n \mapsto
a^{3^n}$  only guarantees the existence of such a number $a$.  How can we
determine this number $a$ and will it have a value, making the function also of
practical interest?
\item Is there an infinite amount of Mersenne prime numbers?
\item Is there an infinite amount of Fermat prime numbers?

@@

\paragraph{Further interesting topics regarding prime numbers}
This chapter doesn't consider the following number theory topics such as
divisibility rules, modulus calculation, modular inverses, modular powers and
roots, Chinese remainder theorem, Euler PHI function, perfect numbers.

@@

@@

So far, no formal proof has been found that perfectly secure cryptographic hash
functions exist. However, there are several good candidates that have not yet
shown any weak points in practice (e.g. \index{SHA--1} SHA--­1 or \index{RIPEMD--
160} RIPEMD--­160).

@@

Mathematicians have been researching elliptic curves for over 100 years. In the
course of time, many lengthy and mathematically complex results have been found
and published in connection with elliptic curves. A mathematician would say that
elliptic curves (or the mathematics behind them) are widely understood. This
research was originally purely mathematical. That is to say, elliptic curves
were investigated, for example, in the mathematical areas of number theory and
algebraic geometry, which are generally highly abstract. Even in the recent
past, elliptic curves played an important role in pure mathematics. In 1993 and
1994, Andrew Wiles published mathematical works that triggered enthusiasm far
beyond the specialist audience. In these works, he proved a conjecture put
forward in the 1960's. To put it short, this conjecture was concerned with the
connection between elliptic curves and what are called module forms. That which
is interesting for most people is that the works of Wiles also proved the famous
second theorem of Fermat. Mathematicians had spent centuries (Fermat lived from
1601 to 1665) trying to find a strict proof of this theorem. Understandably,
therefore, Wiles' proof got a good response. Fermat formulated his theorem as
follows (written in the border of a book):

@@

\paragraph{Cyclic groups}
Cyclic groups are those groups $G'$ that possess an element $g$
from which the group operation can be used to generate all other
elements in the group. This means that for each element $a$ in
$G'$ there exists a positive whole number $i$ such that if $g$ is
subject to the operation $i$ times (i.e. ''$g \cdot i$''),
$g+g+\cdots+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of
the cyclic group --- each element in $G´$ can be generated using
$g$ and the operation.

@@

We can now define an operation\footnote{
A animated addition of elliptic curve points can be found at the web page of Certicom\index{Certicom}:
\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\tt http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}.}
(also denoted as $+$, although it is
not the standard/usual addition of real numbers) on two elements in ${\bf E}$ such that the operation delivers an element
that also lies in ${\bf E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the point in infinity $O$. Thus, for every two
points $(x_1,y_1)$ and $(x_2,y_2)$ on the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such that
the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) = (x_3,y_3)$. Under certain circumstances, these points
may also be equal to the point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf E}$,
we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two points on an elliptic curve specified by $a, b$ and $p$
can therefore be added and the result is a point that also lies on the same elliptic curve.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{ELL1_50.eps}
\caption{Doubling the point P} \vspace{\floatsep}
\includegraphics[scale=0.5]{ELL2_50.eps}
\caption{ Adding distinct points P and Q}
\end{center}
\end{figure}

We must note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation ''addition of $(x_1,y_1)$ and $(x_2,y_2)$'')
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
\end{itemize}
For cryptography, the important fact is that, for very large numbers, it appears to be extremely difficult to use a
given point $Q$ on an elliptic curve to determine which two points have to be added to obtain $Q$.

@@

For secure curves, the parameter $p$ determines how long it takes
to solve the ECLDP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDLP problem.

@@

The {\em elliptic curve discrete logarithm problem} (ECLDP) forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called
{\em domain parameters}.\index{Domain parameters} They determine on which elliptic curve ${\bf
E}$ and in which cyclic subgroup of ${\bf E}$ a signature
procedure has been ''used''.

@@

\newpage
%\addcontentsline{toc}{subsection}{Literaturverzeichnis}
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}
	\bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra\\ \index{Lenstra 1987}
		Factoring integers with elliptic curves, Annals of Mathematics 126, pp. 649-673, 1987.
\end{thebibliography}
\newpage

@@

\begin{enumerate}
   \item Certicom – Online Tutorial, index{Certicom}\\
		\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\tt http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}
	\item IEEE P1363

		\href{http://grouper.ieee.org/groups/1363}{\tt http://grouper.ieee.org/groups/1363}
	\item  \hypertarget{Lenstra2}
An informative web page to the factorisation with elliptic curves.

>>>>

\documentclass[a4paper,11pt]{article}
% Da man keine Unterst"utzung der deutschen Sprache in diesem Fall braucht!!
% \usepackage[german]{babel}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{makeidx}
\usepackage{graphicx}
\makeindex
\hypersetup{colorlinks=true}
\usepackage[flushmargin]{footmisc}

@@

\begin{document}
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}
%\setlength{\footskip}{4cm}
%\renewcommand{\footnoterule}{\parindent0cm\rule{13cm}{.1pt}\vspace{.2cm}}
%Formatierung der Fu"snoten

%space between text and footnote 
\renewcommand\footnoterule{%
  \vspace{2em}%   <-- one line space between text and footnoterule
%\kern-3\p@
  \hrule width .4\columnwidth
 \vspace{4pt}
%kern 2.6\p@
}

%\long\def\@makefntext#1{%
%    \parindent 1em%
 %   \noindent
%    \hbox to 1.8em{\hss\@makefnmark}#1}

@@

You will receive information about the principles of symmetrical and
asymmetrical encryption. A large section of the script is dedicated to the
fascinating topic of prime numbers. Using numerous examples, the elementary number
theory and modular arithmetic are introduced and applied in an exemplary 
manner in the RSA procedure.  After this, one receives insight into the mathematical ideas
behind modern cryptography.

@@

The {\em authors} Bernhard Esslinger, Bartol Filipovic, Henrik Koy and Roger Oyono
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Giessen and Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

@@

\subsubsection{Asymmetric encryption}

@@

Although there are only $92$ natural chemical elements, the number of prime
numbers is unlimited. Even the Greek, \index{Euclid} Euclid\footnote{Euclid,
a Greek mathematician of 4th and 3rd century before Christ. He worked at the
egyptian academy of Alexandria and wrote ''The Elements'', the most well known systematical textbook
of the Greek mathematics.} knew this in the
third century before Christ.
\begin{theorem}[Euclid]\footnote{The common usage of the term does not denote Euclid as the inventor of the theorem rather;
the true inventor is merely not as prominent. The theorem has already been distinguished
and proven in Euclid’s Elements (Book IX, s. 20). The phraseology is remarkable due to 
the fact that the word infinite is not used. The text reads as following
$$
O\acute{\iota}~\pi\varrho\tilde{\omega}\tau o \iota~\grave{\alpha}\varrho\iota\vartheta\mu o\grave{\iota}~
\pi\lambda\varepsilon\acute{\iota}o \upsilon\varsigma~\varepsilon\grave{\iota}\sigma\grave{\iota}~
\pi\alpha\nu\tau\grave{o}\varsigma~\tau o \tilde{\upsilon}~
\pi\varrho o \tau\varepsilon\vartheta\acute{\varepsilon}\nu\tau o \varsigma~
\pi\lambda\acute{\eta}\vartheta\ o \upsilon\varsigma~
\pi\varrho\acute{\omega}\tau\omega\nu~
\grave{\alpha}\varrho\iota\vartheta\mu\tilde{\omega}\nu,
$$
the English translation of which is: the prime numbers are more than any previously existing amount
of prime numbers.}
The sequence of prime numbers does not discontinue.
Therefore, the quantity of prime numbers is infinite.
\end{theorem}
His proof that there is an infinite number of
primes is still considered to be a brilliant mathematical consideration and
conclusion today (proof by contradiction). He assumed that there is only a
finite number of primes and therefore a largest prime number. Based on this
assumption, he drew logical conclusions until he obtained an obvious
contradiction. This meant that something must be wrong. As there were no
mistakes in the chain of conclusions, it could only be the assumption that was
wrong. Therefore, there must be an infinite number of primes!

@@

Here we may mention yet another fact which is initially somewhat surprising. 
Namely, in the prime numbers sequence $p_1, p_2, \cdots,$ gaps between prime numbers can have
an individually determined length $n$. It is undeniable that under the $n$
succession of natural numbers
$$(n+1)!+2,\cdots, (n+1)!+(n+1),
$$
none of them is a prime number since in order, the numbers $2,3,\cdots,(n+1)$  
are comprised respectively as real divisors. 
Nevertheless $n!$ means the product of the first $n$ natural numbers therefore 
$ n!= n*(n-1)*\cdots *2*1.$

@@

The 39th Mersenne prime, called M-39, $$2^{13,466,917}-1,$$ was found at December 6, 2001 –- more exactly,
the verification of this number, found at November 14, 2001 by the Canadian student Michael Cameron,
was successfully finished. This number has about 4 million decimal digits (exactly 4,053,946 digits).
Trying only to print this number $(924947738006701322247758 \cdots 1130073855470256259071)$ in the Financial Times
would need around 200 pages.

@@

  \item {\em Fermat numbers}\footnote{The Fermat prime numbers play a role in circel division.
As proven by Gauss\index{Gauss}, a regular $p$-edge
can only be constructed with the use of a pair of compasses and a ruler, when $p$ is a Fermat prime number.}
\index{Numbers!Fermat numbers}\index{Fermat!Fermat
number}  $F(n) = 2^{2^n} + 1$:\\
As mentioned above, Fermat wrote to Mersenne regarding this assumption.
Surprisingly he would have been able obtain a positive result using the negative
prime number test for $n=5$ based on his small theorem.
$$
\begin{array}{lll}
F(0) = 2^{2^0} + 1  = 2^1 + 1 & = 3 &   \mapsto {\rm ~prime}  \\
F(1) = 2^{2^1} + 1  = 2^2 + 1 & = 5 &   \mapsto {\rm ~prime}  \\
F(2) = 2^{2^2} + 1  = 2^4 + 1 & = 17 &  \mapsto {\rm ~prime}  \\
F(3) = 2^{2^3} + 1  = 2^8 + 1 & = 257 & \mapsto {\rm ~prime}  \\
F(4) = 2^{2^4} + 1  = 2^{16} + 1 &  = 65,537 &  \mapsto {\rm ~prime}  \\
F(5) = 2^{2^5} + 1  = 2^{32} + 1 &  = 4 ,294, 967, 297 = 641 \cdot
6,700,417 &  \mapsto {\rm ~NOT~prime} !
\end{array} 
$$
  \item Carmichael numbers: see above.
  \item Pseudoprime numbers: see above.
  \item Strong pseudoprime numbers: see above.
  \item Idea based on \hyperlink{euclid}{Euclid's proof} (infinite many prime
numbers) $p_1 \cdot p_2 \cdots p_n +1$:

@@

$e_9, \cdots, e_{17}$ are also composite, which means that this formula is not
particularly useful. Comment: However, what is special about all these numbers
is that any pair of them does not have a common factor other than $1$. They are
therefore \index{Prime numbers!relatively} {\em relatively prime }.
  \item $f(n) = n^2 + n + 41$:\\
  This sequence starts off very {\em promisingly},   but is far from being a
proof.

@@

This is the \index{Prime number theorem} \textbf{prime number theorem}. It was
put forward by Legendre (1752-1833) and Gauss\index{Gauss} (1777-1855) but not proved until
over 100 years later.

@@

\begin{itemize}
  \item The German mathematician \index{Goldbach Christian} Christian Goldbach
(1690-1764) conjectured:
    Every even natural number greater than 2 can be represented as the sum of
two prime numbers.
    Computers have verified\footnote{It is generally accepted today, that
the Goldbach Conjecture is true, i.e. valid for all prime numbers.
In 1999, mathematician J\"org Richtstein from the computer sciences institute at the University
of Giessen, studied even numbers up to 400 billion and found no contradictory example (see
\href{http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}{\tt http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}).\index{Richstein 1999}
Nevertheless, this does not provide us with general proof.\\
The fact that despite all efforts, Goldbachs' conjecture has to date not been proven leads
one to believe that: since the pioneer work of the Austrian mathematician Kurt G\"odel is well-known,
not every true mathematical theorem is provable (see
\href{http://www.mathematik.ch/mathematiker/goedel.html}{http://www.mathematik.ch/mathematiker/goedel.html}).
Perhaps Goldbachs' Conjecture was correct, despite that however,
proof will never be found. Conversely, that will presumably also remain unproven.}
the Goldbach conjecture for all even numbers up to
$4*10^{14}$ but no general proof has yet been found\footnote{The English
publisher {\em Faber} and the American publisher {\em Bloomsbury} issued in 2000
the 1992 published book ''Uncle Petros and Goldbach's Conjecture'' by Apostolos Doxiadis.
It's the story of an old maths professor who fails to prove a more than 250 old puzzle. To
stimulate the sell rates the English and American publisher compete a prize of 1 million USD, 
is someone proves the conjecture -- to be published till 2004 by an well-known mathematical journal.\\
Susprisingly only British and American citizens are allowed to participate.\\
The theorem closest to Goldbach's conjecture yet was proven 1966 by Chen Jing-Run in a hard to understand fashion:
Each even integer bigger than $2$ is the sum of one prime and of the product of two primes. E.g.: $20=5+3*5.$\\
Most of the research about the Goldbach conjecture is collected in the book: ''Goldbach Conjecture'', ed. Wang Yuan, 1984, World scientific
Series in Pure Maths, Vol. 4.}.
  \item The German \index{Riemann Bernhard} mathematician Bernhard Riemann
(1826-1866) put forward a formula for the distribution of primes that would
further improve the estimate. However, this has neither been proved nor
contradicted (?) so far.
\end{itemize}

@@

Twin primes are prime numbers whose difference is 2. Examples include 5 and 7 or
101 and 103. Triplet primes, however, only occur once: 3, 5, 7.  For all other
sets of three consecutive uneven numbers, one of them is always divisible by 3
and thus not a prime.
\begin{itemize}
  \item The number of twin primes is an open question: infinite or limited
number?
  The largest twin primes known today are $1,693,965 \cdot 2^{66,443} \pm 1.$
  \item Does a formula exist for calculating the number of twin primes per
interval?
  \item The above proof of the function  $f: N \rightarrow Z$ with $n \mapsto
a^{3^n}$  only guarantees the existence of such a number $a$.  How can we
determine this number $a$ and will it have a value, making the function also of
practical interest?
\item Is there an infinite number of Mersenne prime numbers?
\item Is there an infinite number of Fermat prime numbers?

@@

\paragraph{Further interesting topics regarding prime numbers}
This chapter doesn't consider the other number theory topics such as
divisibility rules, modulus calculation, modular inverses, modular powers and
roots, Chinese remainder theorem, Euler PHI function, perfect numbers.

@@

\subsection*{Thanks} \addcontentsline{toc}{subsection}{Thanks}

I would like to take this opportunity to thank Mr. Henrik Koy and Mr. Roger Oyono for their very constructively proof-reading of this article.

@@

So far, no formal proof has been found that perfectly secure cryptographic hash
functions exist. However, there are several good candidates that have not yet
shown any weak points in practice (e.g. \index{SHA--1} SHA--1 or \index{RIPEMD--160} RIPEMD-­160).

@@

Mathematicians have been researching elliptic curves for over 100 years. In the
course of time, many lengthy and mathematically complex results have been found
and published in connection with elliptic curves. A mathematician would say that
elliptic curves (or the mathematics behind them) are widely understood. This
research was originally purely mathematical. That is to say, elliptic curves
were investigated, for example, in the mathematical areas of number theory and
algebraic geometry, which are generally highly abstract. Even in the recent
past, elliptic curves played an important role in pure mathematics. In 1993 and
1994, Andrew Wiles\index{Wiles} published mathematical works that triggered enthusiasm far
beyond the specialist audience. In these works, he proved a conjecture put
forward in the 1960's. To put it short, this conjecture was concerned with the
connection between elliptic curves and what are called module forms. That which
is interesting for most people is that the works of Wiles also proved the famous
second theorem of Fermat. Mathematicians had spent centuries (Fermat lived from
1601 to 1665) trying to find a strict proof of this theorem. Understandably,
therefore, Wiles' proof got a good response. Fermat formulated his theorem as
follows (written in the border of a book):

@@

\paragraph{Cyclic groups}
Cyclic groups\index{Groups!cyclic} are those groups $G'$ that possess an element $g$
from which the group operation can be used to generate all other
elements in the group. This means that for each element $a$ in
$G'$ there exists a positive whole number $i$ such that if $g$ is
subject to the operation $i$ times (i.e. ''$g \cdot i$''),
$g+g+\cdots+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of
the cyclic group --- each element in $G´$ can be generated using
$g$ and the operation.

@@

We can now define an operation\footnote{A animated addition of elliptic curve points can be
found at the web page of Certicom\index{Certicom}:
\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\tt http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}.}
(also denoted as $+$, although it is
not the standard/usual addition of real numbers) on two elements in ${\bf E}$ such that the operation delivers an element
that also lies in ${\bf E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the point in infinity $O$. Thus, for every two
points $(x_1,y_1)$ and $(x_2,y_2)$ on the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such that
the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) = (x_3,y_3)$. Under certain circumstances, these points
may also be equal to the point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf E}$,
we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two points on an elliptic curve specified by $a, b$ and $p$
can therefore be added and the result is a point that also lies on the same elliptic curve.

\newpage
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{ELL1_50.eps}
\caption{Doubling the point P} \vspace{\floatsep}
\includegraphics[scale=0.5]{ELL2_50.eps}
\caption{ Adding distinct points P and Q\footnotemark}
\end{center}
\end{figure}
\addtocounter{footnote}{0}\footnotetext{The point $O$ cannot be represented in the affine plane.}
\newpage
We must note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation ''addition of $(x_1,y_1)$ and $(x_2,y_2)$'')
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
\end{itemize}
For cryptography, the important fact is that, for very large numbers, it appears to be extremely difficult to use a
given point $Q$ on an elliptic curve to determine which two points have to be added to obtain $Q$.

@@

For secure curves, the parameter $p$ determines how long it takes
to solve the ECDLP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDLP problem.

@@

The {\em elliptic curve discrete logarithm problem} (ECDLP) forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called
{\em domain parameters}.\index{Domain parameters} They determine on which elliptic curve ${\bf
E}$ and in which cyclic subgroup of ${\bf E}$ a signature
procedure has been ''used''.

@@

\newpage
%\addcontentsline{toc}{subsection}{Literaturverzeichnis}
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}
	\bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra\\ \index{Lenstra 1987}
		Factoring integers with elliptic curves, Annals of Mathematics 126, pp. 649-673, 1987.
\end{thebibliography}
%\newpage

@@

\begin{enumerate}
   \item Certicom – Online Tutorial, \index{Certicom}\\
		\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\tt http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}
	\item IEEE P1363

		\href{http://grouper.ieee.org/groups/1363}{\tt http://grouper.ieee.org/groups/1363}
	\item  \hypertarget{Lenstra2}{}
An informative web page to the factorisation with elliptic curves.
